### 2025-01-04 Documents/presentation/sections/1_introduction.tex
presentation

\begin{frame}{Wstęp}
  Czym jest algorithm trading
\end{frame}